A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-23 paper/paper.tex
paper: \filler -> \painting

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\filler}[2][1,2]{\prim{painting}[][#1][#2][][]}
\newcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\newcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\newcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohc}{\coh{painting}[][][][][][][][]}
\newcommandx{\coht}{\cohtwo{frame}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

@@

Our precise mechanization in Coq is available on Github, at \href{https://github.com/artagnon/bonak}{github.com/artagnon/bonak}. The construction was conceived in Summer 2019 and the mechanization began in late 2019. A sketch of the construction was presented at the 2020 HoTT-UF workshop and the completion of the mechanization was reported at the TYPES 2022 conference.

@@

  The definition of $\DeltaPlus$ is shown in Fig. \ref{fig:deltaplus}. Note that, if $g \circ f$ is well-defined, then the length of $f$ is less than that of $g$. It can be shown that composition is associative and that $\id$ is neutral.
\end{definition}

@@

\subsection{Intuition for our formal construction}
To assign types to $X_0, X_1, X_2, \ldots$ in the above indexed representation, we give a recursive definition relying on building blocks, which we call $\framep$, $\layer$, and $\filler$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), which we decompose into $\layer$s, and a $\filler$ corresponds to a filled $\framep$. Some $\framep$ are full and we call them $\fullframe$.

We assign every $X_n$ the type $\fullframe[n] \rightarrow \U$ uniformly, applying to the description above the isomorphism between $A \rightarrow B \rightarrow C$ and $A \times B \rightarrow C$, or between $\Pi a : A, (B a \rightarrow C)$ and $(\Sigma a : A, B a) \rightarrow C$. Let us now illustrate how we build $\fullframe[n]$. To begin, let us set $\fullframe[0] = \unittype$, so that the type $\U$ of $X_0$ can be equivalently formulated as $\unittype \rightarrow \U$. Then, more generally, we let each $\fullframe[n]$ consist of $n$ layers, written $\layer[n][p]$ with $p < n$, that we stack in order, starting from the $\unittype$ type and writing $\framep[n][p]$ for the $p$ first layers of a $\fullframe[n]$, so that $\fullframe[n]$ is $\framep[n][n]$. For instance, $X_1$ is made of one layer, so that it can be written as a $\Sigma$-type of an inhabitant of the $\unittype$ and $\layer[1][0]$, as shown in Fig. \ref{fig:intuition} for $\nu = 2$ which also mentions how the type of $X_2$ is structured.

\begin{figure*}[!t]
  \begin{equation*}
    \begin{array}{ll}
      X_0                                : \underbrace{\unittype}_{\framep[0][0]}                                                    \rightarrow \U \\
      X_1                                : \underbrace{\Sigma \unitpoint: \unittype. \underbrace{\left(
      \begin{array}{c}
          \underbrace{X_0(\unitpoint)}_{\filler[0][0]}
          \\ \times \\
          \underbrace{X_0(\unitpoint)}_{\filler[0][0]}
        \end{array}\right)}_{\layer[1][0]}}_{\framep[1][1]} \rightarrow \U                                                                            \\
      X_2                                : \underbrace{\Sigma a: \underbrace{\left(\Sigma \unitpoint: \unittype. \underbrace{\left(
          \begin{array}{c}
            \underbrace{\Sigma b: \left(
            \hspace{-0.4em}\begin{array}{c}
                             X_0(\unitpoint)
                             \\ \times \\
                             X_0(\unitpoint)
                           \end{array}\hspace{-0.4em}
            \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][L]}}_{\filler[1][1]}}_{\filler[1][0]}
            \\ \times \\
            \underbrace{\Sigma b: \left(
            \hspace{-0.4em}\begin{array}{c}
                             X_0(\unitpoint)
                             \\ \times \\
                             X_0(\unitpoint)
                           \end{array}\hspace{-0.4em}
            \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][R]}}_{\filler[1][1]}}_{\filler[1][0]}
          \end{array}
          \right)}_{\layer[2][0]}\right)}_{\framep[2][1]}.\underbrace{\left(
        \begin{array}{c}
          \underbrace{X_1 \underbrace{\left(a.\hd, \left(
              \hspace{-0.4em}\begin{array}{c}
                               a.\tl.L.\hd.L, \\
                               a.\tl.R.\hd.L
                             \end{array}\hspace{-0.4em}
              \right)\right)}_{\restrf[2][1][L]}}_{\filler[1][1]}
          \\ \times \\
          \underbrace{X_1 \underbrace{\left(a.\hd, \left(
              \hspace{-0.4em}\begin{array}{c}
                               a.\tl.L.\hd.R, \\
                               a.\tl.R.\hd.R
                             \end{array}\hspace{-0.4em}
              \right)\right)}_{\restrf[2][1][R]}}_{\filler[1][1]}
        \end{array}
        \right)}_{\layer[2][1]}}_{\framep[2][2]}
      \rightarrow \U                                                                                                                                \\
      \ldots
    \end{array}
  \end{equation*}
  \caption{Intuition for formal construction ($\nu = 2$)}\label{fig:intuition}
\end{figure*}

@@

We illustrated here the cubical case, that is $\nu = 2$, but, in general, a $\layer[n][p]$ is a product of $\nu$ $\filler[n - 1][p]$. A $\filler[n][0]$ is a $n$-dimensional object corresponding to a filled $\fullframe[n]$. More generally, a $\filler[n][p]$ is an $n$-dimensional object which has the form of a $\filler[n-p][0]$, thus of $(n-p)$-dimensional form, but shifted and living in dimensions $p$ to $n$. Such $\filler[n][p]$ fills a space framed by a partial $\framep[n][p]$ so that, together, they form a filled $\fullframe[n]$. For instance, in the picture, each of the two $\filler[2][0]$ of $\layer[3][0]$ is a filled blue square, each of the two $\filler[2][1]$ of $\layer[3][1]$ is the line shown in red, stretched into a partial square filling the partial frames made of respectively the left and right border of the blue square, and each of the two $\filler[2][2]$ of $\layer[3][2]$ is a point shown in green, stretched into a partial square filling the full frames made respectively of the upper and lower borders of the blue and red squares. A $\filler[n][p]$ adds to a $\framep[n][p]$ the layers needed to form a $\fullframe[n]$. They are added from dimension $n$ to dimension $p$ that is opposite to the order from $0$ to $p$ the $\framep[n][p]$ are built. A $\filler[n][p]$ complementing a given $\framep[n][p]$ fills the resulting $\fullframe[n]$. Thus, on the contrary of frames, it includes a filler, which lives in $X_n$ as shown below:

\begin{align*}
  \framep[n][p] & \defeq \Sigma a_n : (\ldots (\Sigma \unitpoint : \unittype .\; \layer[n][0]) \ldots) .\; \layer[n][p] \\
  \filler[n][p] & \defeq \Sigma l_p : \layer[n][p] . (\ldots (\Sigma l_n : \layer[n][n - 1] .\; X_n) \ldots)
\end{align*}

So far, we have not paid attention to the fact that we have a dependent type, shown as a $\Sigma$. Let us be more precise about this requirement. First, $\fullframe[n]$ depends on all $X_i$ up to $n - 1$. So, we need to package up $X_i$, for $i < n$, into a type, which we call $\X[][<n][]$. This allows us to give the type $\X[][<n][] \rightarrow \U$ to $\fullframe[n]$. Then, for $D: \X[][<n][]$, representing an initial prefix of $X_0, X_1, \ldots X_{n - 1}$, the indexed set $X_n$ has type $\fullframe[n](D) \rightarrow \U$. Thus, $\framep[n][p]$, $\layer[n][p]$ and $\filler[n][p]$ also depend on $D$. We can then refine the previous equation by showing the dependencies on $D$. In particular, $X_n$ is just $D.\tl$.

\begin{align*}
   & \framep[n][p](D) \defeq                                                                                     \\
   & \;\;\;\;\Sigma a_n : (\ldots (\Sigma \unitpoint : \unittype .\; \layer[n][0](D)) \ldots) .\,\layer[n][p](D) \\
   & \filler[n][p](D) \defeq                                                                                     \\
   & \;\;\;\;\Sigma l_p : \layer[n][p](D).\, (\ldots (\Sigma l_n : \layer[n][n - 1](D) .\, D.\tl) \ldots)
\end{align*}

An extra refinement arises from the fact that each addition of a layer to a frame has to be glued onto the border of the partial frame built so far. So, each $\layer[n][p]$ has to depend on $\framep[n][p]$. We also need a way to characterize the border of the $\nu$ $\filler[n-1][p]$ that compose a $\layer[n][p]$, and this is $\restrf[n][p][\epsilon]$, for all $\epsilon < \nu$. For instance, on the picture, each of the $\filler[2][1]$, shown as the red boxes, is laid on the borders of the blue boxes, and hence needs to depend on $\framep[3][1]$. The left and right borders of the two blue boxes are extracted as $\restrf[2][1][L](D)(d)$ and $\restrf[2][1][R](D)(d)$. We can then refine the previous equation by showing the dependencies on $d$, as shown in Figure~\ref{fig:refinement}.

\begin{figure*}[!t]
  \begin{align*}
    \framep[n][p](D)    & \defeq & \Sigma d : (\ldots (\Sigma \unitpoint: \unittype.\, \layer[n][0](D)(\unitpoint)) \ldots).\, \layer[n][p](D)(d)                                \\
    \filler[n][p](D)(d) & \defeq & \Sigma l_p: \layer[n][p](D)(d).\, (\ldots (\Sigma l_n : \layer[n][n - 1](D)(d, l_p, \ldots, l_{n - 1}).\, D.\tl(d, l_p, \ldots, l_n)) \ldots) \\
                        &        & \text{where } (d, l_p, \ldots, l_q) \text{ abbreviates } ((\ldots(d, l_p), \ldots), l_q)
  \end{align*}
  \caption{Refinement of definition of $\framep$ and $\filler$\label{fig:refinement}}
\end{figure*}

Let us recall that a layer is made of $\nu$ paintings. When $\nu=2$, this amounts to:

\begin{align*}
  \layer[n][p](D)(d) & \defeq & \filler[n-1][p](D.\hd)(\restrf[n][p][L][p](d)) & \times \\
                     &        & \filler[n-1][p](D.\hd)(\restrf[n][p][R][p](d)) &
\end{align*}

$\restrf[n][p][\epsilon][q]$ indicates restriction on $q$ layers of a frame, and the induction is on $p$, going from $0$ to $q$. In particular, $\restrf[n][p][\epsilon][p]$ is a ``full restriction''. We define $\restrf[n][p][\epsilon][p](d)$ by recursion on the structure of a frame $d$, which necessitates definitions of $\restrl[n][p][\epsilon][q](d)(l)$ and $\restrc[n][p][\epsilon][q](d)(c)$, for $l$ a $\layer$ and $c$ a $\filler$. The key case is $\restrc[n][p][\epsilon][p](d)(c)$, where $c$, a $\filler[n][p]$, has necessarily the form of $((c_L, c_R), \_)$: $\restrc[n][p][L][p]$ picks out $c_L$, a $\filler[n-1][p]$, $\restrc[n][p][R][p]$ picks out the $c_R$, also a $\filler[n-1][p]$, and $\_$, a $\filler[n][p+1]$, is discarded. There is one more difficulty, which we illustrate by writing down expected and actual types.

@@

\begin{align*}
  c_\omega & : \filler[n-1][p](D.\hd)(\restrf[n-1][p][\omega][q](d))
\end{align*}
$\restrl[n][p][\epsilon](d)(c_L, c_R)$ produces a layer, in which the $\omega$-component has the type

\begin{equation*}
  \filler[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\epsilon][q](\restrf[n][p][\omega][p](d)))
\end{equation*}
while we expect a term of type

\begin{equation*}
  \filler[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\omega][p](\restrf[n][p][\epsilon][q+1](d)))
\end{equation*}

Hence, we need a coherence condition to commute the restrictions. Coherence conditions similar to this necessitate, what are shown as, $\cohf$, $\cohl$ and $\cohc$ in tables in the next section. These are by induction on the structure of $\framep$, $\layer$ and $\filler$. Note that, for the construction in \ref{sec:itt}, we further need a $2$-dimensional coherence condition, $\coht$, for $\cohl$.

@@

% Frame, layer, and painting
% Drop the universe letter
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][#2][#3][#4][#5]}
\renewcommandx{\filler}[5][1,2,3,4,5]{\prim{painting}[][#2][#3][#4][#5]}

@@

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames}, \ref{tab:faces} and \ref{tab:coh}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a coinductive limit of $n$-truncated $\nu$-sets. The $n$-truncated $\nu$-sets are themselves described in \ref{tab:core}. The $n$th component is a family over a $\fullframe$. The type $\fullframe$ is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\layer$ and $\filler$. The type $\layer$ and $\filler$ are dependent over $\restrf$, and these restrictions are defined on \ref{tab:faces}. These restrictions are defined using auxiliary definitions of $\restrl$ and $\restrc$.

Notably, the definition of $\restrl$ relies on a definitional equality expressing the commutation of the composition of $\restrf$. This commutation is not provable by computation so we have to prove it propositionally before using the reflection rule. Proving this itself requires an induction on the dimension, and on the structure of $\framep$, $\layer$, and $\filler$. This is what $\cohf$ proves, as shown in the table \ref{tab:coh}, using auxiliary definitions $\cohl$ and $\cohc$. Even though it looks independent of the other tables, it has to proved mutually with the definitions of $\framep$, $\layer$, $\filler$, and their corresponding restrictions.

@@

% Frame, layer, and painting
% Drop all arguments for coh table; they're not present in coind table.
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][][][][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][][][][#5]}
\renewcommandx{\filler}[5][1,2,3,4,5]{\prim{painting}[][][][][#5]}

@@

For a fixed $n$, $\framep$, $\restrf$, and $\cohf$ have to be mutually defined by induction on $p$. Each of them rely on $\filler$, $\restrf$, and $\cohc$ at lower $n$, so they are well-founded. Finally, each of $\filler$, $\restrf$, and $\cohf$ are by induction from $p$ to $n$.

@@

% Keep all arguments for all macros
% Frame, layer, and painting
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[#1][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[#1][#2][#3][#4][#5]}
\renewcommandx{\filler}[5][1,2,3,4,5]{\prim{painting}[#1][#2][#3][#4][#5]}
\renewcommandx{\fullframe}[3][1,2,3]{\prim{fullframe}[#1][#2][][][#3]}

@@

\subsection{Details on the mechanization\label{sec:mech}}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\filler}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][#1][][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][#1][][]}
\renewcommandx{\cohc}[1][1]{\coh{painting}[][][][][][#1][][]}

>>>>

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\newcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\newcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\newcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohc}{\coh{painting}[][][][][][][][]}
\newcommandx{\coht}{\cohtwo{frame}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

@@

Our precise mechanization in Coq is available at \href{https://github.com/artagnon/bonak}{github.com/artagnon/bonak}. The construction was conceived in Summer 2019 and the mechanization began in late 2019. A sketch of the construction was presented at the 2020 HoTT-UF workshop and the completion of the mechanization was reported at the TYPES 2022 conference.

@@

  The definition of $\DeltaPlus$ is shown in Fig.~\ref{fig:deltaplus}. Note that, if $g \circ f$ is well-defined, then the length of $f$ is less than that of $g$. It can be shown that composition is associative and that $\id$ is neutral.
\end{definition}

@@

\subsection{Intuition for our formal construction}
To assign types to $X_0, X_1, X_2, \ldots$ in the above indexed representation, we give a recursive definition relying on building blocks, which we call $\framep$, $\layer$, and $\painting$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), which we decompose into $\layer$s, and a $\painting$ corresponds to a filled $\framep$. Some $\framep$ are full and we call them $\fullframe$.

We assign every $X_n$ the type $\fullframe[n] \rightarrow \U$ uniformly, applying to the description above the isomorphism between $A \rightarrow B \rightarrow C$ and $A \times B \rightarrow C$, or between $\Pi a : A, (B a \rightarrow C)$ and $(\Sigma a : A, B a) \rightarrow C$. Let us now illustrate how we build $\fullframe[n]$. To begin, let us set $\fullframe[0] = \unittype$, so that the type $\U$ of $X_0$ can be equivalently formulated as $\unittype \rightarrow \U$. Then, more generally, we let each $\fullframe[n]$ consist of $n$ layers, written $\layer[n][p]$ with $p < n$, that we stack in order, starting from the $\unittype$ type and writing $\framep[n][p]$ for the $p$ first layers of a $\fullframe[n]$, so that $\fullframe[n]$ is $\framep[n][n]$. For instance, $X_1$ is made of one layer, so that it can be written as a $\Sigma$-type of an inhabitant of the $\unittype$ and $\layer[1][0]$, as shown in Fig.~\ref{fig:intuition} for $\nu = 2$ which also mentions how the type of $X_2$ is structured.

\begin{figure*}[!t]
  \begin{equation*}
    \begin{array}{ll}
      X_0                                : \underbrace{\unittype}_{\framep[0][0]}                                                    \rightarrow \U \\
      X_1                                : \underbrace{\Sigma \unitpoint: \unittype. \underbrace{\left(
      \begin{array}{c}
          \underbrace{X_0(\unitpoint)}_{\painting[0][0]}
          \\ \times \\
          \underbrace{X_0(\unitpoint)}_{\painting[0][0]}
        \end{array}\right)}_{\layer[1][0]}}_{\framep[1][1]} \rightarrow \U                                                                            \\
      X_2                                : \underbrace{\Sigma a: \underbrace{\left(\Sigma \unitpoint: \unittype. \underbrace{\left(
          \begin{array}{c}
            \underbrace{\Sigma b: \left(
            \hspace{-0.4em}\begin{array}{c}
                             X_0(\unitpoint)
                             \\ \times \\
                             X_0(\unitpoint)
                           \end{array}\hspace{-0.4em}
            \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][L]}}_{\painting[1][1]}}_{\painting[1][0]}
            \\ \times \\
            \underbrace{\Sigma b: \left(
            \hspace{-0.4em}\begin{array}{c}
                             X_0(\unitpoint)
                             \\ \times \\
                             X_0(\unitpoint)
                           \end{array}\hspace{-0.4em}
            \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][R]}}_{\painting[1][1]}}_{\painting[1][0]}
          \end{array}
          \right)}_{\layer[2][0]}\right)}_{\framep[2][1]}.\underbrace{\left(
        \begin{array}{c}
          \underbrace{X_1 \underbrace{\left(a.\hd, \left(
              \hspace{-0.4em}\begin{array}{c}
                               a.\tl.L.\hd.L, \\
                               a.\tl.R.\hd.L
                             \end{array}\hspace{-0.4em}
              \right)\right)}_{\restrf[2][1][L]}}_{\painting[1][1]}
          \\ \times \\
          \underbrace{X_1 \underbrace{\left(a.\hd, \left(
              \hspace{-0.4em}\begin{array}{c}
                               a.\tl.L.\hd.R, \\
                               a.\tl.R.\hd.R
                             \end{array}\hspace{-0.4em}
              \right)\right)}_{\restrf[2][1][R]}}_{\painting[1][1]}
        \end{array}
        \right)}_{\layer[2][1]}}_{\framep[2][2]}
      \rightarrow \U                                                                                                                                \\
      \ldots
    \end{array}
  \end{equation*}
  \caption{Intuition for formal construction ($\nu = 2$)}\label{fig:intuition}
\end{figure*}

@@

We illustrated here the cubical case, that is $\nu = 2$, but, in general, a $\layer[n][p]$ is a product of $\nu$ $\painting[n - 1][p]$. A $\painting[n][0]$ is a $n$-dimensional object corresponding to a filled $\fullframe[n]$. More generally, a $\painting[n][p]$ is an $n$-dimensional object which has the form of a $\painting[n-p][0]$, thus of $(n-p)$-dimensional form, but shifted and living in dimensions $p$ to $n$. Such $\painting[n][p]$ fills a space framed by a partial $\framep[n][p]$ so that, together, they form a filled $\fullframe[n]$. For instance, in the picture, each of the two $\painting[2][0]$ of $\layer[3][0]$ is a filled blue square, each of the two $\painting[2][1]$ of $\layer[3][1]$ is the line shown in red, stretched into a partial square filling the partial frames made of respectively the left and right border of the blue square, and each of the two $\painting[2][2]$ of $\layer[3][2]$ is a point shown in green, stretched into a partial square filling the full frames made respectively of the upper and lower borders of the blue and red squares. A $\painting[n][p]$ adds to a $\framep[n][p]$ the layers needed to form a $\fullframe[n]$. They are added from dimension $n$ to dimension $p$ that is opposite to the order from $0$ to $p$ the $\framep[n][p]$ are built. A $\painting[n][p]$ complementing a given $\framep[n][p]$ fills the resulting $\fullframe[n]$. Thus, on the contrary of frames, it includes a filler, which lives in $X_n$ as shown below:

\begin{align*}
  \framep[n][p]   & \defeq \Sigma a_n : (\ldots (\Sigma \unitpoint : \unittype .\; \layer[n][0]) \ldots) .\; \layer[n][p] \\
  \painting[n][p] & \defeq \Sigma l_p : \layer[n][p] . (\ldots (\Sigma l_n : \layer[n][n - 1] .\; X_n) \ldots)
\end{align*}

So far, we have not paid attention to the fact that we have a dependent type, shown as a $\Sigma$. Let us be more precise about this requirement. First, $\fullframe[n]$ depends on all $X_i$ up to $n - 1$. So, we need to package up $X_i$, for $i < n$, into a type, which we call $\X[][<n][]$. This allows us to give the type $\X[][<n][] \rightarrow \U$ to $\fullframe[n]$. Then, for $D: \X[][<n][]$, representing an initial prefix of $X_0, X_1, \ldots X_{n - 1}$, the indexed set $X_n$ has type $\fullframe[n](D) \rightarrow \U$. Thus, $\framep[n][p]$, $\layer[n][p]$ and $\painting[n][p]$ also depend on $D$. We can then refine the previous equation by showing the dependencies on $D$. In particular, $X_n$ is just $D.\tl$.

\begin{align*}
   & \framep[n][p](D) \defeq                                                                                     \\
   & \;\;\;\;\Sigma a_n : (\ldots (\Sigma \unitpoint : \unittype .\; \layer[n][0](D)) \ldots) .\,\layer[n][p](D) \\
   & \painting[n][p](D) \defeq                                                                                   \\
   & \;\;\;\;\Sigma l_p : \layer[n][p](D).\, (\ldots (\Sigma l_n : \layer[n][n - 1](D) .\, D.\tl) \ldots)
\end{align*}

An extra refinement arises from the fact that each addition of a layer to a frame has to be glued onto the border of the partial frame built so far. So, each $\layer[n][p]$ has to depend on $\framep[n][p]$. We also need a way to characterize the border of the $\nu$ $\painting[n-1][p]$ that compose a $\layer[n][p]$, and this is $\restrf[n][p][\epsilon]$, for all $\epsilon < \nu$. For instance, on the picture, each of the $\painting[2][1]$, shown as the red boxes, is laid on the borders of the blue boxes, and hence needs to depend on $\framep[3][1]$. The left and right borders of the two blue boxes are extracted as $\restrf[2][1][L](D)(d)$ and $\restrf[2][1][R](D)(d)$. We can then refine the previous equation by showing the dependencies on $d$, as shown in Fig.~\ref{fig:refinement}.

\begin{figure*}[!t]
  \begin{align*}
    \framep[n][p](D)      & \defeq & \Sigma d : (\ldots (\Sigma \unitpoint: \unittype.\, \layer[n][0](D)(\unitpoint)) \ldots).\, \layer[n][p](D)(d)                                \\
    \painting[n][p](D)(d) & \defeq & \Sigma l_p: \layer[n][p](D)(d).\, (\ldots (\Sigma l_n : \layer[n][n - 1](D)(d, l_p, \ldots, l_{n - 1}).\, D.\tl(d, l_p, \ldots, l_n)) \ldots) \\
                          &        & \text{where } (d, l_p, \ldots, l_q) \text{ abbreviates } ((\ldots(d, l_p), \ldots), l_q)
  \end{align*}
  \caption{Refinement of definition of $\framep$ and $\painting$\label{fig:refinement}}
\end{figure*}

Let us recall that a layer is made of $\nu$ paintings. When $\nu = 2$, this amounts to:

\begin{align*}
  \layer[n][p](D)(d) & \defeq & \painting[n-1][p](D.\hd)(\restrf[n][p][L][p](d)) & \times \\
                     &        & \painting[n-1][p](D.\hd)(\restrf[n][p][R][p](d)) &
\end{align*}

$\restrf[n][p][\epsilon][q]$ indicates restriction on $q$ layers of a frame, and the induction is on $p$, going from $0$ to $q$. In particular, $\restrf[n][p][\epsilon][p]$ is a ``full restriction''. We define $\restrf[n][p][\epsilon][p](d)$ by recursion on the structure of a frame $d$, which necessitates definitions of $\restrl[n][p][\epsilon][q](d)(l)$ and $\restrc[n][p][\epsilon][q](d)(c)$, for $l$ a $\layer$ and $c$ a $\painting$. The key case is $\restrc[n][p][\epsilon][p](d)(c)$, where $c$, a $\painting[n][p]$, has necessarily the form of $((c_L, c_R), \_)$: $\restrc[n][p][L][p]$ picks out $c_L$, a $\painting[n-1][p]$, $\restrc[n][p][R][p]$ picks out the $c_R$, also a $\painting[n-1][p]$, and $\_$, a $\painting[n][p+1]$, is discarded. There is one more difficulty, which we illustrate by writing down expected and actual types.

@@

\begin{align*}
  c_\omega & : \painting[n-1][p](D.\hd)(\restrf[n-1][p][\omega][q](d))
\end{align*}
$\restrl[n][p][\epsilon](d)(c_L, c_R)$ produces a layer, in which the $\omega$-component has the type

\begin{equation*}
  \painting[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\epsilon][q](\restrf[n][p][\omega][p](d)))
\end{equation*}
while we expect a term of type

\begin{equation*}
  \painting[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\omega][p](\restrf[n][p][\epsilon][q+1](d)))
\end{equation*}

Hence, we need a coherence condition to commute the restrictions. Coherence conditions similar to this necessitate, what are shown as, $\cohf$, $\cohl$ and $\cohc$ in tables in the next section. These are by induction on the structure of $\framep$, $\layer$ and $\painting$. Note that, for the construction in \ref{sec:itt}, we further need a $2$-dimensional coherence condition, $\coht$, for $\cohl$.

@@

% Frame, layer, and painting
% Drop the universe letter
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][#2][#3][#4][#5]}
\renewcommandx{\painting}[5][1,2,3,4,5]{\prim{painting}[][#2][#3][#4][#5]}

@@

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames}, \ref{tab:faces} and \ref{tab:coh}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a coinductive limit of $n$-truncated $\nu$-sets. The $n$-truncated $\nu$-sets are themselves described in \ref{tab:core}. The $n$th component is a family over a $\fullframe$. The type $\fullframe$ is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\layer$ and $\painting$. The type $\layer$ and $\painting$ are dependent over $\restrf$, and these restrictions are defined on \ref{tab:faces}. These restrictions are defined using auxiliary definitions of $\restrl$ and $\restrc$.

Notably, the definition of $\restrl$ relies on a definitional equality expressing the commutation of the composition of $\restrf$. This commutation is not provable by computation so we have to prove it propositionally before using the reflection rule. Proving this itself requires an induction on the dimension, and on the structure of $\framep$, $\layer$, and $\painting$. This is what $\cohf$ proves, as shown in the table \ref{tab:coh}, using auxiliary definitions $\cohl$ and $\cohc$. Even though it looks independent of the other tables, it has to proved mutually with the definitions of $\framep$, $\layer$, $\painting$, and their corresponding restrictions.

@@

% Frame, layer, and painting
% Drop all arguments for coh table; they're not present in coind table.
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][][][][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][][][][#5]}
\renewcommandx{\painting}[5][1,2,3,4,5]{\prim{painting}[][][][][#5]}

@@

For a fixed $n$, $\framep$, $\restrf$, and $\cohf$ have to be mutually defined by induction on $p$. Each of them rely on $\painting$, $\restrf$, and $\cohc$ at lower $n$, so they are well-founded. Finally, each of $\painting$, $\restrf$, and $\cohf$ are by induction from $p$ to $n$.

@@

% Keep all arguments for all macros
% Frame, layer, and painting
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[#1][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[#1][#2][#3][#4][#5]}
\renewcommandx{\painting}[5][1,2,3,4,5]{\prim{painting}[#1][#2][#3][#4][#5]}
\renewcommandx{\fullframe}[3][1,2,3]{\prim{fullframe}[#1][#2][][][#3]}

@@

\subsection{Details on the mechanization\label{sec:mech}}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][#1][][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][#1][][]}
\renewcommandx{\cohc}[1][1]{\coh{painting}[][][][][][#1][][]}

Since the construction shown in the previous sections are by induction on $n$, and dependencies are on lower $n$ and $p < n$, one would imagine formalizing this using well-founded induction in the calculus of inductive constructions. We initially tried this approach: we had terms dependent on the proofs of the case distinction of $n' \leq n$ implies $n' < n$ or $n' = n$, and these proofs didn't have definitional computational rule; it started to be necessary to reason propositionally on the computational property of the case distinction, and it eventually turned out to be unmanageable. Hence, we chose a different route: in practice, since $\restrf[n]$ depends on $\framep[n]$ and $\framep[n-1]$, while $\cohf[n]$ depends on $\framep[n]$, $\framep[n-1]$, and $\framep[n-2]$, we only need to keep track of three ``levels'', and we built separate data structures for the levels, with dependencies. More concretely, we build the ten definitions shown in the tables by induction, and this is part of the definition of a larger record. The other fields of the record aref $\framep$, $\layer$, $\painting$ at levels $n - 1$ and $n - 2$, $\restrf$, $\restrl$, and $\restrc$ at level $n - 1$, and equations to recall the definitions of these objects at lower levels.
